---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/Figures/README-",
  fig.align = "center",
  out.width = "75%",
  cache = TRUE
)
```

# viralprod

<!-- badges: start -->
[![Funding](https://img.shields.io/static/v1?label=powered+by&message=lifewatch.be&labelColor=1a4e8a&color=f15922)](https://lifewatch.be)
[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

`viralprod` is a powerful R package designed for automating viral production data analyses. It is a valuable tool for researchers and marine scientists investigating marine viruses, enabling them to calculate, analyze and visualize viral production efficiently.

You can find detailed information in the vignettes online:

* [Input data for viralprod](https://github.com/mdhishamshaikh/ViralProduction_R/tree/rpkg/vignettes)
* [Introduction to viralprod](https://github.com/mdhishamshaikh/ViralProduction_R/tree/rpkg/vignettes)

## Installation

You can install the development version of `viralprod` from [GitHub](https://github.com/) using the following code:

``` r
# install.packages("devtools")
devtools::install_github("mdhishamshaikh/ViralProduction_R", dependencies = TRUE, build_vignettes = TRUE)
```

To load the package, use the following command:

```{r, results='hide', warning=FALSE, message=FALSE}
library(viralprod)
```

## Getting started

The `viralprod` package simplifies the analysis of viral production data, making it suitable for researchers studying marine viruses. To utilize the package effectively, it's essential to complete prior steps, including sampling, virus reduction assay encompassing filtration and incubation stages, and subsequent flow cytometry processing. The package is structured around **three** fundamental steps, each contributing to a comprehensive analysis:

* *Calculating Viral Production*: It provides two primary methods, `linear regression` and `VIPCAL`, for calculating viral production rates from viral counts. 
* *Analyze Viral Production*: Various parameters for estimating virus-mediated microbial mortality can be determined based on the calculated viral production values.
* *Visualize Viral Production*: Effective data visualization plays a pivotal role in data analyses. The package provides illustrative examples to visualize your viral production data effectively.

Begin by exploring the available functions within the package. Alongside the pipe operator from the `dplyr` package, a variety of functions are at your disposal:

```{r}
ls("package:viralprod")
```

Within the package, each of the three primary steps has its own, dedicated function. Additionally, a comprehensive wrapper function exists, `vp_end_to_end`, which integrates all three essential steps of viral production analyses into a single execution. Executing this singular function yields calculated viral production results, analyzed data, and visualizations. Below, we will demonstrate the functionality of the wrapper function using example data available in the `inst/extdata` folder, use of the separate main functions is integrated in the `Introduction to viralprod` vignette. 

Running the code, provided below, also serves as a convenient method to verify the correct installation of the package. Note that the assessment of input data frames conforms to the package's requirements is seamlessly integrated into the wrapper function through the `vp_class_count_data` function. 

For detailed information regarding the specific requirements for input data, please consult the vignette: `Input data for viralprod`. Furthermore, comprehensive insights into the methodology, including individual function usage, will be expounded upon in `Introduction to viralprod`.

Two input data files are required: count data from flow cytometer, the original abundances of seawater sample. Load in example data:

```{r}
data_NJ2020_all <- read.csv(system.file('extdata', 'NJ2020_Station_2_and_6_all_populations.csv', package = "viralprod"))
head(data_NJ2020_all)
```

```{r}
str(data_NJ2020_all)
```

```{r}
NJ2020_original_abundances <- read.csv(system.file('extdata','NJ2020_original_abundances.csv', package = "viralprod"))
NJ2020_original_abundances
```

```{r}
str(NJ2020_original_abundances)
```

## Package's usage

Before executing the comprehensive wrapper function, let's take a look at all the different arguments:

```{r}
args(vp_end_to_end)
```

In the initial stages of the wrapper function, both input data frames undergo a thorough examination to ensure they conform to the required format. Subsequently, we encounter the `methods` parameter, which governs the selection of algorithmic variants of `linear regression` and `VIPCAL` for the calculation of viral production rates. As a user, you have the flexibility to execute all available variants or opt for a subset, among the prominent methods are:

* Method **4**: vp_linear_average_replicates_diff.
* Method **9**: vp_VIPCAL_average_replicates_diff.
* Method **12**: vp_VIPCAL_average_replicates_diff_LMER_SE.

Furthermore, the wrapper function offers a range of supplementary parameters. In the absence of user-defined values, default settings defined within the function, are automatically applied. Notably, the package incorporates a dedicated function designed to generate a comprehensive list of available variants for viral production calculation in the global environment. 

```{r}
vp_list_of_methods()
names(list_of_methods)
```

For example, consider a closer look at method **12**:

```{r}
list_of_methods[12]
```

Now, it's time to execute the wrapper function. Although the following demonstration uses example data, it's essential to note that executing this function can serve as a convenient method to verify the correct installation of the package. Since this is for illustration purposes, no output files will be written. If `write_output` is set to `TRUE`, ensure to specify the output directory where the results should be stored. All output data frames and a list containing various visualizations will become available in the global environment. 

```{r, warning=FALSE, message=FALSE}
# All methods will be executed with default parameters
vp_end_to_end(data = data_NJ2020_all,
              original_abundances = NJ2020_original_abundances,
              methods = c(1:12),
              SR_calc = TRUE,
              BP_endpoint = TRUE,
              burst_sizes = c(),
              bacterial_secondary_production = NULL,
              nutrient_content_bacteria = list(),
              nutrient_content_viruses = list(),
              write_output = FALSE,
              output_dir = "")
```

After running the wrapper function, check the global environment to access its output:

```{r}
ls()
```

The `calculation` step provides by default four data frames:

* *vp_results_output_df* contains the viral production results for all samples.
* *vp_results_output_T24_df* contains the viral production results for all samples at the end of the assay.
* *vp_results_output_SR_df* contains the viral production results of the separate replicate treatment, no averaging over replicates.
* *vp_results_output_BP_df* contains the viral production results for all samples with the bacterial endpoint taken into account. 

```{r}
head(vp_results_output_df)
```

The `analyze` step enhances the viral production results by adding various parameters to estimate virus-mediated microbial mortality. An additional data frame provides descriptions and units for each variable.

```{r}
str(analyzed_vp_results_df)
```

```{r}
analyzed_vp_results_dictionary
```

The `visualize` step generates a list of plot objects. The package offers eight different ways to visualize the original viral count data, calculated viral production data, or analyzed viral production data.

```{r}
names(plot_list)
```

For instance, let's start with an overview of the count data of Station 2 for each sample type. The bacterial endpoint for this assay is highlighted in pink. It's worth noting that stopping the assay at this point is advisable to prevent an increase in collision rates in VP samples, which may result in higher-than-expected lytic viral production.

```{r}
grid::grid.draw(plot_list[[1]]$plot_object)
```

Additionally, the package provides multiple ways for calculating viral production. While `linear regression` uses the slope between the count data, `VIPCAL` looks at the average of increments to determine viral production. `VIPCAL-SE` takes it a step further by considering standard errors, ensuring that only true increments are used for calculation. As shown in the figures, `VIPCAL` may overestimate viral production, while `VIPCAL-SE` tends to be more conservative. 

```{r}
plot_list[[6]]$plot_object
```

```{r}
plot_list[[9]]$plot_object
```

One of the analyzed variables considers the nutrient release from bacteria and viruses, mapping the total nutrient release in the sample at the end of the assay. 

```{r}
plot_list[[11]]$plot_object
```

Now that you have explored the available functions and the wrapper function, you are ready to harness the full potential of the `viralprod` package for your own viral production data analysis. A more detailed and comprehensive tutorial on the package's functionality and usage, can be found in the `Introduction to viralprod` vignette. This vignette provides in-depth insights into various underlying methods and features of the package, enabling you to leverage the package's capabilities effectively.  

## License

MIT. See `LICENSE.md`

This software is developed for scientific, educational and research purposes. It is not meant to be used for legal or economical purposes.
